\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[final,expansion=false]{microtype}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{3em}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem*{theoremtarget}{Theorem Target}
\newtheorem*{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\forcesH}{\Vdash_{H}}
\newcommand{\nforcesH}{\nVdash_{H}}
\newcommand{\forcesnn}{\Vdash_{\neg\neg}}
\newcommand{\nforcesnn}{\nVdash_{\neg\neg}}
\newcommand{\PH}{\mathbb{P}_{H}}
\newcommand{\PHeta}{\mathbb{P}_{H}^{\eta}}
\newcommand{\refines}{\leq}
\newcommand{\Gam}{\tilde{\Gamma}}
\newcommand{\lmin}{\lambda_{\min}}
\newcommand{\Adm}{\mathrm{Adm}}
\newcommand{\Stab}{\mathrm{Stab}}
\newcommand{\Kt}{K_{t}}
\newcommand{\Dadm}{\mathcal{D}_{\mathrm{adm}}}
\newcommand{\Jadm}{J_{\mathrm{adm}}}
\newcommand{\Jwit}{J_{\mathrm{wit}}}
\newcommand{\Jnn}{J_{\neg\neg}}
\newcommand{\Sspec}{\mathcal{S}_{\mathrm{spec}}}

\title{Forcing Correspondence (v16):\\ Calibrated Structural Transduction Between\\
Set-Theoretic Forcing and the Hypothesis Surface}
\author{Paul Tiffany}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We give a structural correspondence between set-theoretic (Cohen) forcing and the refinement
 dynamics of a resource-bounded hypothesis surface. Geometric forcing is \emph{not} identical to
 Cohen forcing; the claim is a conditional transduction theorem. Both instantiate a common
 \emph{forcing-over-a-site} schema: the forcing relation is \emph{dense stabilization}---a condition
 forces a proposition when the refinements stabilizing it are dense below it in the channel-margin
 subposet---and the operationally admissible dense requirements \emph{generate} a Grothendieck
 topology $\Jadm$ on the channel-margin subposet, sandwiched below the double-negation topology, so
 the Kripke--Joyal interpretation transfers clause by clause. The site is the closure of obligation
 under refinement: raw admissible obligations $\Dadm$ generate the semantic site $\Jadm$, of which a
 finite-prefix \emph{witnessed} fragment $\Jwit\subseteq\Jadm$ is the executable discipline.
 Channel-margin preservation is a path-budget perturbation lemma (the per-step form does not
 iterate); separativity is carried by certificate disagreement, not by the metric margin; generic existence is Rasiowa--Sikorski under the observer
 bounded-countability condition; the generic is read geometrically as a Cauchy-generic completion in
 an observer completion, with an order--metric compatibility lemma tying the generic filter to the
 limit state, conditional on a smooth descent contract coextensive with the channel margin. Atomic
 soundness follows from eventual stabilization; propositional soundness follows by the standard
 forcing induction, conditional on a decision-reachability lemma and a calibration hypothesis here
 decomposed into a finite theorem queue rather than asserted. Inter-observer transfer is restored as
 an essential interface medium $M$: the void supplies commensurability; a public medium and protocol
 transport only encoded residues $E_A(T_Ax_A)$, and reconstruction is generically not identity---an
 equality-condition theorem. The executable interpretation is witness-only PyLantern/Chalked traces.
 Quantifiers, fullness/witness clauses, and implementation calibration remain open; every load-bearing
 claim is tagged P/D/S/M/C/O in a status ledger (\S\ref{sec:ledger}).
\end{abstract}

\section{Two senses of ``forcing''}

The word is currently overloaded across this program; separating the two uses is notation, not a
retraction.

\emph{Control forcing} (Hypothesis Surface \S4.1) is the precondition: the geometric cost of
joint progress on $k$ claims diverges at the conflict threshold, so a bounded observer is forced,
dynamically, to stop differentiating and integrate. This is the event that deposits the observer
onto a refinement poset.

\emph{Logical forcing} ($\forcesH$, this note) is the relation that then lives on that poset. We
reserve ``forces'' for the second and ``forced transition'' for the first.

\section{Core definitions (instantiated)}

\begin{definition}[Conditions]\label{def:cond}
$P_H$ is the set of claim complexes $\Kt$ with fibers $f_i=(\sigma_i,r_i,W_i)$ and exit cone
$E(x_0)$. A condition is a finite partial epistemic state.
\end{definition}

\begin{assumption}[Smoothness contract]\label{asm:smooth}
Fix a margin $\eta>0$ and an anchor condition $p_0$ with $\lmin(\Gam(p_0))\ge\eta$. Along
admissible trajectories from $p_0$: (a)~$\Gam$ is $L_\Gamma$-Lipschitz in spectral norm with
respect to the projected state displacement; (b)~the cumulative displacement obeys the budget
$\sum_n\|\Delta_n\|\le\eta/(2L_\Gamma)$. This is the displacement contract of Cacophony \S2.1 in
path form; it is \emph{measured, not assumed} (see the remark following the Channel-Margin Lemma
in \S\ref{sec:target}), and it is the single calibration debt written ``(ii)'' in earlier versions
of this note.
\end{assumption}

\begin{definition}[Refinement order and channel-margin subposet]\label{def:refine}
$q\refines p$ iff $q$ is reachable from $p$ by an admissible move retaining $p$'s commitments. The
\emph{channel-margin subposet} $\PHeta$ consists of the conditions reachable from the anchor $p_0$
by admissible moves within the displacement budget of Assumption~\ref{asm:smooth}. All forcing
below is relative to $\PHeta$. (That $\lmin(\Gam)\ge\eta/2$ holds throughout $\PHeta$ is proved,
not stipulated, by the Channel-Margin Lemma of \S\ref{sec:target}; the definition consumes only the budget.)
\end{definition}

\begin{definition}[Sieves and dense requirements]\label{def:req}
For $p\in\PHeta$ a \emph{sieve} on $p$ is a downward-closed subset $S\subseteq{\downarrow}p=\{q\in\PHeta:q\refines p\}$.
A family $D$ of conditions below $p$ is an \emph{admissible dense
requirement} if for every $q\refines p$ in $\PHeta$ there is a further \emph{admissible} refinement
$r\refines q$ in $\PHeta$ meeting $D$ (the paradigm cases are certificate transitions and voids,
but the class is not restricted to them---deciding families must be eligible); the sieve it generates is
$S_D=\{q\refines p:\exists\,\text{admissible }r\refines q,\ r\in D\}$. Only admissible dense
requirements count; not every void or certificate the observer names qualifies. Write $\Dadm$ for the
class of all admissible dense requirements on $\PHeta$.
\end{definition}

\begin{definition}[Stability, atomic]\label{def:stab}
For \emph{atomic} $\varphi$ in the enrolled vocabulary, $\Stab(q,\varphi)$ holds when $\varphi$
holds at $q$ and persists under further admissible refinement (the soft-min propagation of Eq.~1
makes confidence monotone down the order; that Eq.~1 propagates this way is a modeling postulate
about the apparatus---the highest-leverage postulate in this note, consumed by \emph{every} truth
lemma below, tracked as (M-Pers) in \S\ref{sec:ledger}). The \emph{stability set} of atomic
$\varphi$ below $p$ is
\[
  A_\varphi(p)=\{\,q\refines p:\Stab(q,\varphi)\,\}.
\]
By (M-Pers), $A_\varphi(p)$ is closed under further refinement, hence a sieve on $p$. $\Stab$ is
\emph{not} defined for compound formulas; compound forcing is defined recursively by the
propositional clauses of \S\ref{sec:schema}, and the torsion remark of \S\ref{sec:target} records
where a compound-$\Stab$ reading would diverge from them.
\end{definition}

\begin{definition}[Preservation set]
$\mathcal{I}=\{$I1 No Bare Claims (well-definedness), I2 Anti-Masking (boundedness: keeps $\Gam$
faithful, hence $\lmin$ honest), I3 Certificate Classification (obstruction typing)$\}$.
\end{definition}

\section{The generated admissible site: raw, generated, witnessed}\label{sec:site}

The forcing-over-refinement schema needs a site. We do \emph{not} claim that the raw operational
class $\Dadm$ is already closed under the Grothendieck operations---raw admissibility depends on
channel margin, smoothness, finite witnessability, resource bounds, and protocol constraints, and may
fail pullback/transitivity closure as an operational class. Instead we \emph{generate} the topology.

\begin{definition}[Generated admissible site]\label{def:site}
Let $B=\{S_D:D\in\Dadm\}$ be the collection of sieves generated by admissible dense requirements
(Def.~\ref{def:req}). Let
\[
  \Jadm=\langle B\rangle
\]
be the least Grothendieck topology on the poset $\PHeta$ containing $B$. The pair $(\PHeta,\Jadm)$ is
the \emph{generated admissible site}.
\end{definition}


\begin{lemma}[Site bound]\label{lem:sitebound}
$\Jadm$ exists and is a Grothendieck topology, and $\Jadm\subseteq\Jnn$, where $\Jnn$ is the
double-negation (dense) topology whose covers are exactly the dense sieves. In particular
$(\PHeta,\Jadm)$ is a genuine site.
\end{lemma}

\begin{proof}
The least Grothendieck topology containing an arbitrary collection of sieves always exists: the
intersection of all Grothendieck topologies containing $B$ is again a Grothendieck topology and is
the least such. For the bound, recall that on any poset the dense sieves form a Grothendieck topology,
the double-negation topology $\Jnn$ (a sieve $S$ on $p$ is $\Jnn$-covering iff for every $q\refines p$
some $r\refines q$ lies in $S$; see \cite{maclane-moerdijk}). Each generator $S_D\in B$ is dense below
$p$ by admissible density (Def.~\ref{def:req}), hence $S_D\in\Jnn$. Since $\Jnn$ is a Grothendieck
topology containing $B$, the least such is contained in it: $\Jadm\subseteq\Jnn$.
\end{proof}

\begin{definition}[Forcing relations]\label{def:force}
For $p\in\PHeta$ and atomic $\varphi$ (stability sets $A_\varphi(p)$, Def.~\ref{def:stab}):
\[
  p \forcesH \varphi \;:\Longleftrightarrow\; A_\varphi(p)\in\Jadm(p),
  \qquad\qquad
  p \forcesnn \varphi \;:\Longleftrightarrow\; A_\varphi(p)\ \text{is dense below }p\ \text{(a }\Jnn\text{-cover)}.
\]
$\forcesH$ is the \emph{calibrated} relation of this note; $\forcesnn$ is the ambient dense-below
reading, the characterization of Cohen forcing transported onto $\Kt$. By
Lemma~\ref{lem:sitebound}, $p\forcesH\varphi\Rightarrow p\forcesnn\varphi$. The two relations are
distinguished notationally throughout; each lemma below states which relation it is about, and
transfer between them is never silent (it is exactly the Decision Reachability question). Both
extend to compound formulas by the recursive clauses of \S\ref{sec:schema}. The strictly stronger
condition ``$\Stab(q,\varphi)$ for \emph{every} $q\refines p$'' (the maximal sieve, a cover in
every topology) forces under both relations and is reserved for torsion (\S\ref{sec:target},
Remark).
\end{definition}

\begin{remark}[The site is the closure of obligation under refinement]
Rigor does not flatten admissibility; it completes it. The operational layer supplies admissible
dense \emph{obligations} $\Dadm$; the mathematical layer closes them under the Grothendieck operations
into the site $(\PHeta,\Jadm)$. Kripke--Joyal forcing over $(\PHeta,\Jadm)$ supplies the abstract
forcing semantics (\S\ref{sec:schema}). $\forcesnn$ is the Kripke--Joyal reading for the ambient
$\Jnn$; the calibrated relation $\forcesH$ uses the generated subtopology $\Jadm\subseteq\Jnn$,
hence is \emph{at least as strong}. Whether a given deciding sieve is actually a $\Jadm$-cover is
the Decision Reachability question (Lemma~\ref{lem:reach}).
\end{remark}

\begin{definition}[Witnessed fragment]\label{def:wit}
The \emph{witnessed fragment} $\Jwit\subseteq\Jadm$ is the sub-collection of $\Jadm$-covers that admit
a finite admissible protocol witness: a finite prefix of refinements, each carrying a certificate,
admissibility check, provenance, and (where required) a human-gated decision, whose density below $p$
is verified by that finite data. $\Jwit$ is in general \emph{not} a topology; it is the executable
basis. The three layers
\[
  \Dadm \;\leadsto\; \Jadm \;\leadsto\; \Jwit\subseteq\Jadm
\]
read: raw operational obligations $\leadsto$ generated semantic closure $\leadsto$ finite-prefix,
PyLantern/Chalked-witnessable fragment.
\end{definition}

\begin{remark}[Self / other / global]
The three layers are the self/other/global stack. Raw is what the self emits; generated is what the
other may require; witnessed is what the global surface can audit. A hypothesis leaves the self as a
raw obligation, meets the other as generated refinement pressure, and stabilizes globally only as
witnessed residue. Formal closure must not silently authorize operational traversal: \emph{the
generated site supplies semantics; the witnessed fragment supplies executable discipline}
($\Jadm\supseteq\Jwit$ is a containment, not an equality, and \S\ref{sec:pylantern} operates only in
$\Jwit$). This is exactly why Chalked (\S\ref{sec:chalked}) matters: it is the finite witness surface
on which $\Jwit$ becomes consent-gated and inspectable.
\end{remark}

\section{Correspondence table}

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}p{0.21\textwidth} p{0.30\textwidth} p{0.41\textwidth}@{}}
\toprule
\textbf{Component} & \textbf{Cohen / set-theoretic} & \textbf{Hypothesis Surface (concrete)} \\
\midrule
Condition & Finite partial function & Claim complex $\Kt$ + fibers; exit cone $E(x_0)$ \\
Refinement $q\refines p$ & Stronger condition & Admissible move in $\PHeta$ \\
Channel viability & (nondegeneracy of extensions) & $\lmin(\Gam)>0$ (off-cliff; $\rho(k-1)<1$) \\
Separativity / incompatibility & Incompatible extensions below $p$ & \emph{Certificate disagreement} (I3); undecided obstruction (Lem~\ref{lem:nonatom}) \\
Requirement / cover & Dense set of conditions & Admissible dense requirement (Def.~\ref{def:req}); $\Jadm$-cover (Def.~\ref{def:site}) \\
Generic / completion & Generic filter $G$ & Admissible-complete trajectory in $\PHeta$ \\
Forcing $p\Vdash\varphi$ & Forcing-set dense below $p$ & $A_\varphi(p)$ a $\Jadm$-cover of $p$ ($\forcesH$); dense form is $\forcesnn$ \\
Extension & $M[G]$ & Stabilized surface along the completed trajectory \\
Preservation & Ground theory preserved & Invariants I1--I3 survive completion \\
Forced negation & $p\Vdash\neg\varphi$ & Torsion certificate (atomic $\varphi$ refuted on \emph{all} refinements; \S\ref{sec:target}) \\
Undecided & $p\nVdash\varphi,\ p\nVdash\neg\varphi$ & Non-torsion certificate (contingent) \\
Soundness & Truth Lemma & Atomic + Propositional Truth Lemma (Thm.~\ref{thm:prop}); quantifiers open \\
Shared zero / origin & Top condition $\mathbb{1}$ & Void (Axiom H.0); unique shared depth-$0$ condition \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Separativity correction.} The metric margin and separativity are distinct roles:
\emph{margin keeps the channel open; certificates make branches incompatible.} The convex metric part
of every exit cone is self-compatible (Lemma~\ref{lem:nonatom}, using convexity), so $\lmin>0$
supplies channel \emph{viability}---the existence of admissible refinements---not separativity.
Incompatibility, hence the separativity needed for a nondegenerate forcing poset, is carried by the
discrete certificate axis (I3): two refinements are incompatible exactly when they disagree on an
undecided obstruction.

\section{Formal schema}\label{sec:schema}

\begin{quote}
\textbf{Forcing schema.} $p\forcesH\varphi$ when the refinements of $p$ that stabilize $\varphi$
form a $\Jadm$-cover of $p$ in the channel-margin subposet.
\end{quote}

\begin{remark}[Rigorous home]
The schema is the persistence-monotone (intuitionistic) reading of forcing on a site: Kripke--Joyal
forcing semantics over $(\PHeta,\Jadm)$, with admissible dense requirements generating the covering
families. Cohen forcing is the set-theoretic instance; its topos presentation routes through
Kripke--Joyal semantics and the double-negation/classicalization construction \cite{maclane-moerdijk},
i.e.\ through $\Jnn$, of which $\Jadm$ is the generated subtopology (Lemma~\ref{lem:sitebound}).
``Calibrated structural transduction'' is then the claim that both Cohen forcing and the Hypothesis
Surface instantiate forcing-over-a-site, the calibration being which requirements generate the site,
and the preserved invariant being the Kripke--Joyal interpretation of the connectives. The transfer is
verifiable clause by clause.
\end{remark}

\begin{definition}[Propositional clauses and deciding sets]\label{def:clauses}
Let $J\in\{\Jadm,\Jnn\}$ and write $\Vdash$ for the corresponding atomic relation of
Def.~\ref{def:force}. Take $\neg,\wedge,\vee$ primitive over $\PHeta$ and define
$\varphi\to\psi:=\neg\varphi\vee\psi$:
\begin{align*}
p\Vdash\varphi\wedge\psi &\iff p\Vdash\varphi \text{ and } p\Vdash\psi;\\
p\Vdash\varphi\vee\psi &\iff \{q\refines p: q\Vdash\varphi \text{ or } q\Vdash\psi\}\in J(p);\\
p\Vdash\neg\varphi &\iff \text{no } q\refines p \text{ has } q\Vdash\varphi.
\end{align*}
The last is the universal-refinement clause; at the void it is the across-every-branch statement of
\S\ref{sec:void}. The \emph{deciding set} of $\varphi$ is
$D_\varphi=\{q: q\Vdash\varphi \text{ or } q\Vdash\neg\varphi\}$ (relation as per context;
$\forcesH$ unless marked).
\end{definition}

\begin{lemma}[Persistence and consistency]\label{lem:pers}
For $\Vdash\in\{\forcesH,\forcesnn\}$ and every propositional $\varphi$: if $p\Vdash\varphi$ and
$q\refines p$ then $q\Vdash\varphi$; and no $p$ forces both $\varphi$ and $\neg\varphi$.
\end{lemma}
\begin{proof}
Persistence by induction on $\varphi$: atomic because $A_\varphi(q)=A_\varphi(p)\cap{\downarrow}q$
is the pullback of the sieve $A_\varphi(p)$ along $q\refines p$, and covers pull back in any
Grothendieck topology (for $\Jnn$: dense restricts to dense); $\wedge$ componentwise; $\vee$
because a $J$-cover of $p$ restricts to a $J$-cover of any $q\refines p$ (stability axiom); $\neg$
because its defining cone is downward-closed. Consistency: $p\Vdash\neg\varphi$ says no
$q\refines p$ forces $\varphi$, yet $p\refines p$, so $p\not\Vdash\varphi$. Beyond the
well-formedness of the atomic relation ($A_\varphi$ a sieve, which is (M-Pers),
Def.~\ref{def:stab}), clause-level persistence consumes no additional postulate---it is pullback
stability of the topology; (M-Pers) does further substantive work only in the Atomic Truth Lemma.
\end{proof}

\begin{lemma}[Deciding sets are order-dense]\label{lem:dec}
For $\Vdash\in\{\forcesH,\forcesnn\}$ and every propositional $\varphi$, the deciding set
$D_\varphi$ is dense in the order.
\end{lemma}
\begin{proof}
Given $p$: if some $q\refines p$ forces $\varphi$ then $q\in D_\varphi$; otherwise no $q\refines p$
forces $\varphi$, which is exactly $p\Vdash\neg\varphi$, so $p\in D_\varphi$. Density is the
$\neg$-clause read contrapositively---uniform in $\varphi$ and in the choice of relation.
\end{proof}

\begin{remark}
Order-density of $D_\varphi$ is free; whether $D_\varphi$ is \emph{admissibly} dense---a
$\Jadm$-cover---is not, and is precisely the Decision Reachability question of \S\ref{sec:target}. By
Lemma~\ref{lem:pers}, $D_\varphi$ is closed under refinement, hence a sieve on any $p$ when
restricted below it.
\end{remark}

\section{Calibration as a finite theorem queue}\label{sec:calibration}

The forcing lemmas below are theorems about the abstract site semantics. Their bearing on the
Hypothesis Surface pipeline requires a calibration hypothesis: the concrete surface operations must
realize the Kripke--Joyal clauses they are claimed to instantiate. Calibration is not one foggy
assumption; it is a \emph{finite theorem queue}, each item with a definite home.

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}r p{0.36\textwidth} p{0.42\textwidth}@{}}
\toprule
\# & \textbf{Calibration lemma} & \textbf{Home / status}\\
\midrule
1 & Atomic Stabilization & claim stabilization $=\Stab$; \emph{Hypothesis Surface}\\
2 & Refinement Audit & admissible move $=$ refinement, audit-visible; \emph{PyLantern}\\
3 & Dense Requirement & operational requirements admissibly dense; measurable (C, Assumption~\ref{asm:smooth})\\
4 & Certificate Conjunction & certificate combination $=\wedge$-clause; \emph{Hypothesis Surface}\\
5 & Admissible-Cover Disjunction & branching $=$ dense-cover $\vee$-clause; \emph{Hypothesis Surface}\\
6 & Torsion Negation & torsion certificate $=$ universal-refinement $\neg$; discharged for \emph{atomic} $\varphi$ (\S\ref{sec:atomic}, Remark); compound case \textbf{O} (finite countermodel)\\
7 & Soft-Min Persistence & Eq.~1 $=$ monotone persistence; \emph{Hypothesis Surface} (M-Pers)\\
8 & Decision Reachability & deciding sieve is a $\Jadm$-cover; Lem~\ref{lem:reach}, conditional on Assumption~\ref{asm:smooth}\\
9 & Observer Completion & $(P_O,d_O)$ metric, completion exists; \emph{Principia}\\
10 & Interface Reconstruction & $E_A,D_B$ carry witness/provenance/consent, $D_B\!\circ\! E_A\neq\mathrm{id}$; \emph{PyLantern/Chalked} + Thm~\ref{thm:nonid}\\
\bottomrule
\end{tabular}
\end{center}

Items 1, 4, 5, 7 are supplied by \emph{The Hypothesis Surface}; 2, 10 by PyLantern/Chalked
implementation; 3, 8 reduce to the measurable smoothness contract (Assumption~\ref{asm:smooth}); 6
is discharged here for atomic formulas only; 9 is an interface to \emph{Principia}. \emph{None is a
fresh unknown.} The Propositional Truth Lemma
is operationally meaningful under this queue; discharging it is more urgent than adding quantifiers,
because it is the check that turns structural analogy into verified transduction for the actual
pipeline.

\begin{assumption}[Calibration queue]\label{asm:calibration}
The realization lemmas 1--10 of the table above hold for the deployed pipeline: the concrete
surface operations realize the Kripke--Joyal clauses they are claimed to instantiate. Items already
discharged or reduced retain their homes as listed; the assumption is the conjunction of the
remainder. Status \textbf{O} overall; consumed explicitly wherever a result is claimed \emph{about
the pipeline} rather than about the site semantics.
\end{assumption}

\section{Theorem target, with proofs of the joints}\label{sec:target}

\begin{theoremtarget}
Let $H$ be a hypothesis surface carried by a resource-bounded observer, with claim complex $\Kt$,
interaction matrix $\Gam_t=G_tG_t^{\top}$, and channel projector $T_t$. Suppose: (1)
$\lmin(\Gam_t(p))\ge\eta>0$ at the anchor; (2) admissible trajectories satisfy the smoothness
contract (Assumption~\ref{asm:smooth}), so the Channel-Margin Lemma
(Lemma~\ref{lem:margin}) applies; (3) $\mathcal{D}_p$ is a countable family of admissible dense requirements
generating (with the deciding closure) $\Jadm$-covers, with decision reachability
(Lemma~\ref{lem:reach}); (4) the observer-completion axioms supply Cauchy convergence for admissible
refinement towers (Lemma~\ref{lem:cauchy}) and order--metric compatibility (Lemma~\ref{lem:ordmet}).
Then there is a Cauchy-generic trajectory $G$ through $p$ meeting every requirement and converging to
$x_G\in\overline{P}_O$; along $G$, $\forcesH$ is monotone, $\mathcal{I}$ is preserved, atomic forced
propositions are exactly those eventually stable (Lemma~\ref{lem:atomic}), and soundness extends to
the propositional connectives (Thm.~\ref{thm:prop}) under the calibration queue.
\end{theoremtarget}

\begin{enumerate}[leftmargin=1.6em]
\item \textbf{Monotonicity (persistence).} $p\forcesH\varphi\Rightarrow q\forcesH\varphi$ for
  $q\refines p$, by pullback stability of covers (Lemma~\ref{lem:pers}); no apparatus postulate is
  consumed.
\item \textbf{Channel viability, persistence, branching (Lemmas~\ref{lem:margin}, \ref{lem:nonatom}).}
  $\lmin(\Gam)>0$ is the metric nondegeneracy making admissible refinements exist (viability); the
  Channel-Margin Lemma keeps the margin positive; separativity/branching is carried by certificate
  disagreement (Lemma~\ref{lem:nonatom}), not the convex metric.
\item \textbf{Generic existence (Lemma~\ref{lem:bdd} + Rasiowa--Sikorski).}
\item \textbf{Decision reachability (Lemma~\ref{lem:reach}).}
\item \textbf{Atomic soundness (Lemma~\ref{lem:atomic}); propositional soundness (Thm.~\ref{thm:prop}).}
  Open layers: quantifiers, and the calibration queue (\S\ref{sec:calibration}).
\end{enumerate}

\begin{lemma}[Channel-Margin, path form]\label{lem:margin}
Let $\lambda(q)=\lmin(\Gam(q))$ and let $q$ be reachable from the anchor $p_0$ by an admissible
trajectory with steps $\Delta_1,\dots,\Delta_n$. Under Assumption~\ref{asm:smooth}
($\lambda(p_0)\ge\eta$; $L_\Gamma$-Lipschitz; budget $\sum_i\|\Delta_i\|\le\eta/(2L_\Gamma)$),
\[
  \lambda(q)\;\ge\;\lambda(p_0)-L_\Gamma\sum_{i\le n}\|\Delta_i\|\;\ge\;\eta/2\;>\;0 .
\]
Hence $\PHeta$ (Def.~\ref{def:refine}) stays in the channel regime, uniformly in the depth of the
descent.
\end{lemma}

\begin{proof}
$\lmin$ is $1$-Lipschitz in the spectral norm (Weyl): $|\lambda(q)-\lambda(p_0)|\le\|\Gam(q)-\Gam(p_0)\|$.
By the Lipschitz hypothesis and the triangle inequality along the trajectory,
$\|\Gam(q)-\Gam(p_0)\|\le L_\Gamma\sum_i\|\Delta_i\|\le\eta/2$, so
$\lambda(q)\ge\eta-\eta/2=\eta/2$.
\end{proof}

\begin{remark}[Why the budget, not a step bound]
The v14 form of this lemma imposed only the per-step bound $\|\Delta\|\le\eta/(2L_\Gamma)$ and
claimed the margin ``inductively along the whole selected descent.'' That induction is unsound: from
$\lambda(q_1)\ge\eta/2$ a second step of the same size yields only $\lambda(q_2)\ge0$; per-step
control lets the margin decay linearly in depth. Depth-uniform margin needs the \emph{cumulative}
budget of Assumption~\ref{asm:smooth}(b)---which is exactly the Lyapunov summability that
Lemma~\ref{lem:cauchy} independently requires for the Cauchy-generic completion. Margin preservation
and generic convergence are funded by the same contract; the budget is spent once. Both directions
are machine-checked in the accompanying verification suite (\texttt{Margin.lean}): the path form is
proved, and the per-step form is formally refuted by the linear-decay countermodel
$\lambda_i=4-2i$.
\end{remark}

\begin{remark}[The smoothness hypothesis is substrate-relative and measurable]
Assumption~\ref{asm:smooth}---Lipschitz continuity of $\Gam$ with a summable displacement
budget---is the smooth-regime condition of the displacement contract (Cacophony \S2.1),
$\|x_{t+1}-x_t\|\le L$, and is measured, not assumed: under audit-observer measurement it holds on a
small fraction of trajectories and fails (pivots) on the rest. An observer with continuous internal
state realizes it natively; a stateless sampler realizes it only on the smooth fraction and
otherwise recovers the descent through external staging, re-imposing the budget as explicit
checkpoints. The quantity governing how close to the cliff one may operate is the perturbation
sensitivity $\sim 1/\lmin$ of Cacophony's Thm.~M.20.
\end{remark}

\begin{lemma}[Non-atomicity is undecidedness]\label{lem:nonatom}
Assume $p$ is in the channel-margin regime, so admissible refinements of $p$ exist and fit the
margin budget (the min-norm feasible displacement $\sqrt{\alpha^{\top}\Gam^{-1}\alpha}$ is finite
exactly when $\Gam\succ0$), and assume incompatibility is generated by certificate disagreement (the
metric part of every exit cone being convex, hence self-compatible---postulate M-Cvx). Then $p$ is
\emph{non-atomic}---there are admissible $q_1,q_2\refines p$ with no common admissible lower
bound---iff some obstruction $\omega$ below $p$ is \emph{undecided for the dense reading}:
$p\nforcesnn\varphi_\omega$ and $p\nforcesnn\neg\varphi_\omega$. The lemma is stated for
$\forcesnn$, where ``not forcing'' unpacks densely.
\end{lemma}

\begin{proof}
$(\Leftarrow)$ $p\nforcesnn\neg\varphi_\omega$ unpacks, by the $\neg$-clause, to: some $q_1\refines
p$ has $q_1\forcesnn\varphi_\omega$. And $p\nforcesnn\varphi_\omega$ means $A_{\varphi_\omega}(p)$
is not dense below $p$: some $q_2\refines p$ has no $\varphi_\omega$-stabilizer below it, whence no
$r\refines q_2$ forces $\varphi_\omega$ ($A_{\varphi_\omega}(r)$ is empty and empty sieves are not
dense), i.e.\ $q_2\forcesnn\neg\varphi_\omega$. A common admissible lower bound $r$ would, by
persistence (Lemma~\ref{lem:pers}), force both $\varphi_\omega$ and $\neg\varphi_\omega$,
contradicting consistency; so $q_1\perp q_2$ and $p$ is non-atomic. $(\Rightarrow)$ If every
obstruction below $p$ is decided, any two admissible refinements agree with $p$ on every certificate
classification (persistence); by M-Cvx their only remaining relation is metric and convex, so they
admit a common admissible continuation; $p$ is atomic. Contrapositively, non-atomic implies some
obstruction is undecided.
\end{proof}

\begin{remark}[Where the generic adds information]
Atoms are exactly the decided leaves; branching, hence nontrivial extension, occurs precisely over
undecided obstructions. Metric nondegeneracy ($\lmin>0$) is necessary (viability) but the branching
is carried by the discrete certificate axis (I3): the convex exit cone is by itself atomic. This is
the sharp form of the v7 caution that ``$\lmin>0$ does not by itself prove incompatible extensions
exist''; the missing ingredient is undecidedness, and with it the implication is an equivalence. The
generic adds information over, and only over, the undecided region---the same region the
deparameterization residue of \S\ref{sec:deparam} occupies. The undecidedness criterion is stated
for the dense reading $\forcesnn$; it transfers to the calibrated $\forcesH$ exactly where Decision
Reachability (Lemma~\ref{lem:reach}) makes the two readings agree on deciding sieves.
\end{remark}

\begin{lemma}[Boundedness yields countability for Rasiowa--Sikorski]\label{lem:bdd}
$\mathcal{D}_p$ is countable, the atomic propositions it tracks form a countable vocabulary, and a
filter $G$ through $p$ meeting every admissible dense requirement exists.
\end{lemma}

\begin{proof}
That a resource-bounded observer holds finitely many active claims per stage and at most countably
many over a run is the observer-bound postulate (M-Bound); it supplies only \emph{countability}, not
density. Once each $D\in\mathcal{D}_p$ is verified admissibly dense (Def.~\ref{def:req}), its
generated sieve is a $\Jadm$-cover, in particular order-dense (Lemma~\ref{lem:sitebound}). The
Rasiowa--Sikorski construction applied to this countable family of dense sets yields the filter $G$
\cite{rasiowa-sikorski,gpst}.
\end{proof}

\begin{lemma}[Decision Reachability, conditional]\label{lem:reach}
Assume the smoothness contract (Assumption~\ref{asm:smooth}). Then for every $p\in\PHeta$ and
propositional formula $\varphi$ there is an admissible refinement $q\refines p$, inside the
channel-margin subposet, deciding $\varphi$ ($q\forcesH\varphi$ or $q\forcesH\neg\varphi$);
equivalently, the deciding sieve $D_\varphi\cap{\downarrow}p$ is a $\Jadm$-cover of $p$. Where the
contract fails, order-density need not be admissibly realizable and $D_\varphi$ need not be a
$\Jadm$-cover.
\end{lemma}

\begin{proof}
$D_\varphi$ is dense in the order (Lemma~\ref{lem:dec}) and closed under refinement
(Lemma~\ref{lem:pers}), hence a sieve below any $p$. Under Assumption~\ref{asm:smooth} the
refinements witnessing order-density are realized by margin-respecting admissible moves within the
budget, so $D_\varphi$ is an admissible dense requirement in the sense of Def.~\ref{def:req} (the
definition places no type restriction beyond admissible density). Its generated sieve therefore lies
in the generating family $B$ of Def.~\ref{def:site}, so $D_\varphi\cap{\downarrow}p\in\Jadm(p)$.
The conditionality is irreducible: the step from order-density to admissible density \emph{is} the
contract, not a consequence of it---this is calibration item~8, and the finite model checker of the
verification suite exhibits a site where it fails.
\end{proof}

\begin{remark}[Why this is not free]
Standard forcing gives order-density of deciding refinements (Lemma~\ref{lem:dec}); the Surface needs
\emph{admissible} density inside the channel-margin contract. Decision Reachability names exactly the
missing step: that a deciding refinement is reachable without leaving the channel. It is item~8 of the
calibration queue and is coextensive with the smooth regime, tying bivalence to
Assumption~\ref{asm:smooth} rather than to countability alone. Bivalence
(Lemma~\ref{lem:bivalence}) is therefore P \emph{consuming C}, not P consuming only countability.
\end{remark}

\begin{definition}[Admissible deciding closure]\label{def:closure}
Given a countable admissible dense family $\mathcal{D}_p$, its \emph{admissible deciding closure}
$\mathcal{D}_p^\ast$ adds the atomic stability sets $A_\varphi(p)$ (Def.~\ref{def:stab}) and the
propositional deciding sets $D_\varphi$ (Def.~\ref{def:clauses}), retaining only those that are
$\Jadm$-covers (Lemma~\ref{lem:reach}). Because the observer vocabulary is countable (Lemma~\ref{lem:bdd})
and the propositional closure of a countable vocabulary is countable, $\mathcal{D}_p^\ast$ is still
countable.
\end{definition}

\begin{lemma}[Cauchy--Forcing Completion]\label{lem:cauchy}
Let $(p_n)$ be the admissible chain generating a channel-margin generic $G$ through $p$, with states
$z_n=\Phi(p_n)$ and steps $\Delta_n=z_{n+1}-z_n$. Suppose (a) a residual-instability potential
$V:P_H\to[0,\infty)$ is nonincreasing along admissible refinement; (b) the descent inequality
$V(p_n)-V(p_{n+1})\ge\kappa\|\Delta_n\|$ holds for some $\kappa>0$; and (c) the observer metric
$d_O$ is comparable to the embedding norm along the trajectory. Then $\sum_n\|\Delta_n\|<\infty$,
so $(z_n)$ is Cauchy in the observer metric and converges to $x_G\in\overline{P}_O$.
\end{lemma}

\begin{proof}
By (a), $V(p_n)\downarrow V_\infty\ge0$, so $\sum_n(V(p_n)-V(p_{n+1}))=V(p_0)-V_\infty<\infty$.
By (b), $\|\Delta_n\|\le\kappa^{-1}(V(p_n)-V(p_{n+1}))$, whence $\sum_n\|\Delta_n\|<\infty$.
Absolutely summable increments are Cauchy; by (c), $(z_n)$ is $d_O$-Cauchy and converges in the
completion $\overline{P}_O$.
\end{proof}

\begin{lemma}[Order--Metric Compatibility, atomic]\label{lem:ordmet}
For \emph{atomic} $\varphi$ (the only formulas $\Stab$ is defined for, Def.~\ref{def:stab}), define
two satisfaction relations: $G\models_{\mathrm{ord}}\varphi$ iff the generic filter $G$ meets
the stabilizer requirement $A_\varphi=\{q:\Stab(q,\varphi)\}$; and $x_G\models_{\mathrm{met}}\varphi$ iff the
Cauchy limit $x_G\in\overline{P}_O$ lies in the stabilized region of $\varphi$. Suppose (M-Pers)
persistence, the Cauchy descent of Lemma~\ref{lem:cauchy}, and that the stabilized region
$\{x:\Stab_x(\varphi)\}$ is $d_O$-closed (postulate M-Cl). Then
$G\models_{\mathrm{ord}}\varphi\Rightarrow x_G\models_{\mathrm{met}}\varphi$.
\end{lemma}

\begin{proof}
$G\models_{\mathrm{ord}}\varphi$ gives $q_N\in G$ with $\Stab(q_N,\varphi)$; by (M-Pers)
$\Stab(q_n,\varphi)$ for all $n\ge N$ along the chain, so $z_n$ lies in the stabilized region for
$n\ge N$. By Lemma~\ref{lem:cauchy}, $z_n\to x_G$ in $d_O$; by (M-Cl) the stabilized region is closed,
so its limit $x_G$ lies in it, i.e.\ $x_G\models_{\mathrm{met}}\varphi$.
\end{proof}

\begin{remark}[Convergence is coextensive with the smooth regime]
Hypothesis (b) is the displacement contract (Assumption~\ref{asm:smooth}) in Lyapunov form; its
summability is the same budget that funds the Channel-Margin Lemma. Where the contract holds the descent is
absolutely summable and the generic converges; in the pivot regime (b) fails and the completion can
diverge.
 The convergence of the Cauchy-generic is coextensive with the smooth regime---the same
boundary marked by the channel margin, the deparameterization floor (\S\ref{sec:deparam}), and the
reconstruction floor. The Cauchy debt is that standing hypothesis; completion exists exactly where the
contract holds. Lemma~\ref{lem:ordmet} adds one interface postulate (M-Cl) tying the order-generic to
the metric limit, without which ``$G$ meets requirements'' and ``$x_G$ realizes them'' are distinct
statements.
\end{remark}

\begin{lemma}[Atomic Truth Lemma]\label{lem:atomic}
Let $G$ be an admissible-complete trajectory through $p$ meeting every $\Jadm$-cover in
$\mathcal{D}_p^\ast$, and define $G\models_H\varphi$ iff $\Stab(q,\varphi)$ for some $q\in G$. Then for
atomic $\varphi$ in the observer's countable vocabulary,
$G\models_H\varphi \iff \exists\,p\in G:\ p\forcesH\varphi$.
\end{lemma}

\begin{proof}
$(\Leftarrow)$ If $p\forcesH\varphi$ then $A_\varphi(p)$ is a $\Jadm$-cover (Def.~\ref{def:force});
being a cover, it is included in $\mathcal{D}_p^\ast$ (Def.~\ref{def:closure}), which $G$ meets, so
some $q\in G$ has $\Stab(q,\varphi)$. $(\Rightarrow)$ If $G\models_H\varphi$, some $q\in G$
stabilizes $\varphi$; by (M-Pers), $\Stab(r,\varphi)$ holds for \emph{every} $r\refines q$, so
$A_\varphi(q)={\downarrow}q$ is the maximal sieve on $q$, which belongs to every Grothendieck
topology---in particular to $\Jadm$. Hence $q\forcesH\varphi$ with $q\in G$. (The v14 text
concluded from density here; density does not imply $\Jadm$-covering---the maximal sieve does.)
\end{proof}

\begin{remark}[Certificate typing is forcing-theoretic---atomic case only]\label{rem:torsion}
I3 classifies each impossibility as torsion (structural, permanent) or non-torsion (contingent).
For \emph{atomic} $\varphi$, a torsion certificate at $p$ says $\varphi$ is refuted at every
$q\refines p$; then no $q$ can stabilize $\varphi$, $A_\varphi(q)=\emptyset$ for all $q\refines p$,
no refinement forces $\varphi$ under either relation, and $p\forcesH\neg\varphi$ by the
$\neg$-clause---permanence needing no generic. Conversely, $p\forcesH\neg\varphi$ for atomic
$\varphi$ means no $q$ forces $\varphi$; since a single stabilizing condition forces (its maximal
sieve covers), no $q\refines p$ satisfies $\varphi$ stably. This discharges calibration item~6
\emph{for atomic formulas}. For compound $\varphi$ the identification fails: ``no refinement forces
$\varphi$'' and ``$\neg\varphi$ classically stable everywhere below'' come apart (minimal
countermodel: $\varphi=\neg\beta$ with $\beta$ stabilized exactly at the leaves of a $7$-condition
spine model; no condition forces $\neg\beta$, yet $\neg\neg\beta$ is nowhere classically stable at
interior nodes---see the accompanying verification suite). The compound case is \textbf{O}. In the
Kripke--Joyal reading the split marks where the double-negation construction can later collapse an
undecided condition.
\end{remark}

\section{Propositional soundness: the connective Truth Lemma}\label{sec:atomic}

The remaining joint---extending soundness past atomic propositions---closes at the propositional level,
inside the frame of \S\ref{sec:void}: the obstructing clauses are the universal-refinement ones, which
bottom out at the void. The clauses, persistence, and order-density of deciding sets were established
in \S\ref{sec:schema} (Def.~\ref{def:clauses}, Lemmas~\ref{lem:pers}, \ref{lem:dec}); this section
works with the calibrated relation $\forcesH$ throughout.

\begin{lemma}[Genericity decides every formula]\label{lem:bivalence}
Assume Decision Reachability (Lemma~\ref{lem:reach}). Then a generic $G$ meeting the admissible
deciding closure $\mathcal{D}_p^\ast$ decides every propositional $\varphi$: for each $\varphi$
there is $q\in G$ with $q\forcesH\varphi$ or $q\forcesH\neg\varphi$.
\end{lemma}
\begin{proof}
The propositional closure of the countable atomic vocabulary (Lemma~\ref{lem:bdd}) is countable, so
$\{D_\varphi\}_\varphi$ is countable. Under Lemma~\ref{lem:reach} each $D_\varphi$ restricts to a
$\Jadm$-cover, so each belongs to $\mathcal{D}_p^\ast$ (Def.~\ref{def:closure}), which is countable;
$G$ meets every member, in particular every $D_\varphi$. \emph{Bivalence on $G$ is thus P
conditional on Decision Reachability (C), not on countability alone.}
\end{proof}

\begin{theorem}[Propositional Truth Lemma, conditional on calibration]\label{thm:prop}
Assume the calibration queue (Assumption~\ref{asm:calibration}; in particular Decision
Reachability). Let $G$ be a generic deciding every propositional formula, truth defined classically on
$G$ (bivalence by Lemma~\ref{lem:bivalence}) and atomically by eventual stabilization. Then, under
(C), for every propositional $\varphi$,
$G\models_H\varphi \iff \exists\,p\in G:\ p\forcesH\varphi$.
The forcing-theoretic content is unconditional; (C) is what ties that content to the pipeline, and it
is the antecedent, not a consequence.
\end{theorem}
\begin{proof}
Induction on $\varphi$. \emph{Atomic}: Lemma~\ref{lem:atomic}. \emph{$\wedge$}: $(\Leftarrow)$ a
forcing $p$ gives both conjuncts by the IH; $(\Rightarrow)$ witnesses $p_1,p_2$ have a common lower
bound $p\in G$ (filter), forcing both by persistence. \emph{$\vee$}: $(\Rightarrow)$
$G\models\varphi$ or $\psi$; the IH gives a witness $p\in G$ forcing a disjunct; then the
$\vee$-cover of $p$ contains all of ${\downarrow}p$ by persistence, i.e.\ is the maximal sieve, so
$p\forcesH\varphi\vee\psi$. $(\Leftarrow)$ Let $p\in G$ force $\varphi\vee\psi$ and suppose, for
contradiction, $G\not\models\varphi$ and $G\not\models\psi$. $G$ decides both
(Lemma~\ref{lem:bivalence}), so witnesses $q_1,q_2\in G$ force $\neg\varphi,\neg\psi$; a common
lower bound $s\in G$ of $p,q_1,q_2$ forces $\varphi\vee\psi$, $\neg\varphi$, and $\neg\psi$
(persistence). The $\vee$-cover of $s$ is a $\Jadm$-cover, hence dense
(Lemma~\ref{lem:sitebound}), hence nonempty below $s$: some $r\refines s$ forces $\varphi$ or
$\psi$, contradicting consistency with the persisted $\neg\varphi,\neg\psi$ at $r$
(Lemma~\ref{lem:pers}). (The v14 text appealed to $G$ meeting the $\vee$-cover ``cofinally'';
genericity only guarantees meeting members of $\mathcal{D}_p^\ast$, so the argument now routes
through deciding sets.) \emph{$\neg$}: $(\Leftarrow)$ if $p\in G$ forces $\neg\varphi$ but
$G\models\varphi$, a witness $q\in G$ forces $\varphi$ (IH); a common lower bound $r\in G$ forces both,
contradicting Lemma~\ref{lem:pers}. $(\Rightarrow)$ $G$ decides $\varphi$ (Lemma~\ref{lem:bivalence});
$G\not\models\varphi$ excludes a witness, so the deciding witness forces $\neg\varphi$. Case $\to$
follows from $\neg,\vee$.
\end{proof}

\begin{remark}[Intuitionistic relation, classical extension]
$G$ is classical (bivalent) though the clauses are the intuitionistic (Kripke) ones. This is the
Cohen phenomenon: an intuitionistic forcing relation with a classical generic extension, mediated by
the generic meeting every $\neg\neg$-dense deciding set. The classicality of $G$ is the surface-level
shadow of ``$M[G]$ models the classical theory,'' the operational content of the double-negation
construction of \S\ref{sec:schema}.
\end{remark}

\section{Inter-observer forcing: the shared zeroth-order zero}\label{sec:void}

Single-observer forcing descends within one surface. Two observers with surfaces $\PH^A,\PH^B$
\emph{disjoint below the top}---sharing no condition except the void---raise a separate question: on
what can they agree, when neither can occupy the other's conditions?

\medskip\noindent\textbf{Construction (wedge at the void).} Glue the surfaces at their common top,
\[
  R \;=\; \bigl(\PH^A \sqcup \PH^B\bigr)\big/\bigl(\mathbb{1}_A=\mathbb{1}_B=\mathbb{1}\bigr),
\]
where $\mathbb{1}$ is the void (Axiom H.0): zero commitments, the maximum condition every refinement
descends from. The order on $R$ is the disjoint union of the two branch orders together with
$x\refines\mathbb{1}$ for every $x$; there are \emph{no} cross-branch relations. Each branch keeps
its own order; $\mathbb{1}$ sits above both.

\begin{lemma}[The void is the unique shared point]\label{lem:void}
For $p\in\PH^A\setminus\{\mathbb{1}\}$ and $q\in\PH^B\setminus\{\mathbb{1}\}$, $p$ and $q$ are
incompatible in $R$. Hence $\mathbb{1}$ is the unique condition compatible with both branches, and
any filter in $R$ lies entirely in one branch (plus $\mathbb{1}$).
\end{lemma}

\begin{proof}
A common lower bound $r$ of $p$ and $q$ would satisfy $r\refines p$ and $r\refines q$, and
$r\ne\mathbb{1}$ since $\mathbb{1}$ is above every condition, not below any. By the definition of
the order on $R$ (no cross-branch relations), $r\refines p$ places $r$ in branch $A$ and
$r\refines q$ places $r$ in branch $B$; but $\PH^A\cap\PH^B=\{\mathbb{1}\}$ in $R$. So no common
lower bound exists. A filter is downward-directed, so it cannot contain non-void conditions from
both branches.
\end{proof}

An observer may stand at the origin but descends into only one surface: observation is the choice of
branch, and the two cannot be occupied at once.

\begin{proposition}[The shared point is forced to zeroth order]\label{prop:zeroth}
Order conditions in $R$ by refinement depth from the top. The only condition shareable between
disjoint observers is the unique depth-$0$ condition $\mathbb{1}$---``shared zero point'' is a
consequence, not a choice.
\end{proposition}

\begin{proof}
$\mathbb{1}$ is the unique condition reached by \emph{zero} admissible moves; every other condition
is at depth $\ge1$ and, by Lemma~\ref{lem:void}, branch-private, hence occupiable by at most one of
the two observers. \end{proof}

\noindent(The \emph{order} fact is Proposition~\ref{prop:zeroth}; the identification of that stratum
with the diagnostic zeroth-order register $(\hat\rho,\mu)$, and below with the loopback address, is
a modeling bridge, not a theorem.)

\begin{remark}[Forced from the void]
By Def.~\ref{def:force}, $\mathbb{1}\forcesH\varphi$ iff $\Stab(\cdot,\varphi)$ covers $\mathbb{1}$ in
\emph{both} branches---the branch-invariant propositions. Disjoint observers force the same atomic
$\varphi$ iff it is forced from the void. They agree exactly on the zeroth-order, pre-observational
kernel and on nothing contingent. \emph{The shared zero buys commensurability---a common $\mathbb{1}$
to quantify over, hence the ability to state ``we agree''---not communion.} It transports no contingent
claim across.
\end{remark}

\section{Interface layer: medium, protocol, and the non-identity theorem}\label{sec:interface}

The void is the unique shared depth-$0$ condition. It supplies commensurability, not communication.
Communication requires a first-order medium: a public surface on which one observer exports a
floor-stable residue and another reconstructs it inside its own surface. Networking is therefore
essential to inter-observer work, but its formal role is interface/transport, not shared occupancy:
\emph{void is shared orientation; medium is shared inscription; protocol is shared constraint.}

Let $X_A,X_B$ be the observers' state spaces, carrying the channel projectors $T_A,T_B$ (states
relate to conditions through the embedding $\Phi$ of Lemma~\ref{lem:cauchy}; in v14 the encoding was
typed on the condition poset and applied to states---a category slip, corrected here). Let $M$ be an
interface medium with partial encoding and decoding maps
\[
  E_A: T_A(X_A)\rightharpoonup M,\qquad D_B: M\rightharpoonup X_B .
\]
Observer $A$ exports not the ceiling-state $x_A$ but its floor-stable residue $r_A=T_Ax_A$, encoded
as $m=E_A(r_A)$; $B$ receives a reconstruction $D_B(m)$ inside $B$'s surface.

\begin{definition}[Transfer]\label{def:transfer}
A \emph{transfer of contingent content} from $A$ to $B$ is any correspondence between depth-$\ge1$
content of $A$'s surface and depth-$\ge1$ content of $B$'s that is realized either by shared
occupancy of a condition or by a composite of maps through third objects.
\end{definition}

\begin{theorem}[Factoring, architectural]\label{thm:factor}
If $\PH^A,\PH^B$ are disjoint below the void, then no contingent (depth $\ge 1$) content is shared by
occupancy (Lemma~\ref{lem:void}); hence any transfer of contingent content from $A$ to $B$
(Def.~\ref{def:transfer}) is not occupancy but an external composite
\[
  T_A(X_A) \xrightarrow{\ E_A\ } M \xrightarrow{\ D_B\ } X_B ,
\]
through a third object $M$ that both may address.
\end{theorem}

\begin{proof}
Immediate from Lemma~\ref{lem:void} and Def.~\ref{def:transfer}: the occupancy disjunct is excluded
below the void, leaving the third-object composite. (The theorem is architectural---its content is
the exclusion, carried by Lemma~\ref{lem:void}; it is tagged D, not P, in the ledger.)
\end{proof}

\begin{theorem}[Non-identity of transport, with equality condition]\label{thm:nonid}
Let $T_A$ be the channel projector ($T_A^2=T_A$). For $x_A$ with
$\epsilon_{T_A}(x_A)=\|(\mathrm{id}-T_A)x_A\|/\|x_A\|>0$,
\[
  T_Ax_A\neq x_A,
\]
so the medium is \emph{not} where identity first fails---bounded projection already breaks it.
Consequently $D_B(E_A(T_Ax_A))=x_A$ can hold only under a \emph{named identity-transport condition}:
\begin{enumerate}[leftmargin=1.8em,label=(\roman*)]
\item \textbf{zero projection loss:} $x_A\in\operatorname{im}(T_A)$ (equivalently $\epsilon_{T_A}(x_A)=0$);
\item \textbf{faithful encoding:} $E_A$ injective on the residue $r_A=T_Ax_A$;
\item \textbf{calibrated decoding:} $D_B\circ E_A=\mathrm{id}$ on the residue subobject (a
  section/retraction, i.e.\ an isomorphism of the transported subobjects of $\PH^A$ and $\PH^B$).
\end{enumerate}
Absent any of (i)--(iii), the received state is a reconstruction, not identity.
\end{theorem}

\begin{proof}
$T_Ax_A=x_A\iff x_A\in\ker(\mathrm{id}-T_A)=\operatorname{im}(T_A)\iff\epsilon_{T_A}(x_A)=0$, giving
the first claim and the necessity of (i). If (i) holds, $x_A=r_A$ and
$D_B(E_A(x_A))=x_A$ requires $E_A$ to lose no information on $r_A$ (necessity of (ii)) and
$D_B\circ E_A$ to return $r_A$ within $B$'s surface (necessity of (iii)); failure of either yields a
distinct element. Conversely (i)--(iii) give $D_B(E_A(T_Ax_A))=D_B(E_A(x_A))=x_A$.
\end{proof}

\begin{remark}
Non-identity is generic; identity is a calibrated degeneracy. This is central for Chalked
(\S\ref{sec:chalked}): the board carries residues, not ceilings; a reader gets $D_B(m)$, a
reconstruction, never the author's private state. The invariant is concise: \emph{the void gives
commensurability; the network gives communication}---and communication is reconstruction under $B$'s
projection, priors, and constraints.
\end{remark}

\begin{proposition}[Observer closure is zero projection residue]\label{prop:zombie}
Identify the closure defect of an external observer-closure frame with the projection residue
$\epsilon_T(x)=\|(\mathrm{id}-T)x\|/\|x\|$ of the channel projector $T$ (a modeling bridge, M-Zomb).
Then observer closure ($\epsilon_T=0$) is exactly condition~(i) of Thm.~\ref{thm:nonid}
($x\in\operatorname{im}T$); the behavioral trace transports only the exported floor residue $Tx$, so a
nonzero defect lies in $(\mathrm{id}-T)x$ and is invisible to any finite behavioral test; and by
Prop.~\ref{prop:chi} the certifying identity $\chi_T^2+\epsilon_T^2=1$ holds only under orthogonality,
so a local closure certificate need not globalize. A behaviorally competent system with $\epsilon_T>0$
is a \emph{zombie}; its defect is typed torsion/non-torsion by the decided/undecided axis of
Lem.~\ref{lem:nonatom} and Rem.~\ref{rem:torsion}.
\end{proposition}
\begin{proof}
The three clauses are Thm.~\ref{thm:nonid}(i), the exported-residue definition $r=Tx$
(\S\ref{sec:interface}), and Prop.~\ref{prop:chi}; the typing is Rem.~\ref{rem:torsion}. Only the
identification M-Zomb is new.
\end{proof}

\begin{remark}[Loopback as icon, not transport]
The loopback address \texttt{127.0.0.1} remains useful iconography for indexical self-reference: the
address whose form is universal but whose referent is local. It illustrates the zeroth-order
self-interface; it does not bridge observers. Inter-observer transport requires a non-loopback edge:
a public medium, address, protocol, artifact, or utterance. (Appendix~\ref{sec:appendix}.)
\end{remark}

\section{Interpretive appendix: boundary projection and exportability}\label{sec:eureka}

Lemma~\ref{lem:cauchy} has an observer-side reading we call the \emph{Eureka Tub}: a working term for
the pre-artifact state of insight, not a literal object. A bounded observer momentarily holds a
high-dimensional \emph{ceiling} state---richer than it can carry into record. Exiting imposes a
projection $\pi_O$ onto the \emph{floor}: the residue that can be written, tested, or witnessed. We
identify $\pi_O$ with the channel projector $T_t$ of \S2 (a modeling bridge, M). Define
\[
  \chi_T(x)=\frac{\|T_t x\|}{\|x\|},\qquad
  \epsilon_T(x)=\frac{\|(\mathrm{id}-T_t)x\|}{\|x\|}.
\]

\begin{proposition}[Exportability identity, under orthogonality]\label{prop:chi}
If $T_t$ is an orthogonal projection ($T_t=T_t^2=T_t^{*}$), then for $x\neq0$,
\[
  \chi_T(x)^2+\epsilon_T(x)^2=1 .
\]
If $T_t$ is merely idempotent (oblique), only the triangle bound
$\chi_T(x)+\epsilon_T(x)\ge1$ survives (hence $\chi_T^2+\epsilon_T^2\ge\tfrac12$); the Pythagorean
identity fails in general, with the defect controlled by the principal angle between
$\operatorname{im}T_t$ and $\ker T_t$. Orthogonality must not be smuggled in.
\end{proposition}

\begin{proof}
Orthogonal case: $T_tx\perp(\mathrm{id}-T_t)x$ gives
$\|x\|^2=\|T_tx\|^2+\|(\mathrm{id}-T_t)x\|^2$; divide by $\|x\|^2$. Oblique case:
$x=T_tx+(\mathrm{id}-T_t)x$ and the triangle inequality give
$\|x\|\le\|T_tx\|+\|(\mathrm{id}-T_t)x\|$, i.e.\ $\chi_T+\epsilon_T\ge1$; Cauchy--Schwarz on the
pair $(\chi_T,\epsilon_T)$ gives the quadratic form. That the identity fails obliquely is witnessed
by any idempotent with $\|T_t\|>1$.
\end{proof}

\begin{proposition}[Non-determination of the ceiling-state]\label{prop:skeleton}
Suppose the ceiling space has dimension exceeding the bounded rank of $\pi_O$ (finitely many active
claims). Then $\pi_O$ is non-injective, so no single residue determines the ceiling-state, and
recovery is possible only as the limit of residues along an increasing filtration
$\pi_{O,1}\preceq\pi_{O,2}\preceq\cdots$ whose images are Cauchy---the observer reading of
Lemma~\ref{lem:cauchy}.
\end{proposition}

\begin{proof}
A linear map of rank $r$ on a space of dimension $>r$ has nontrivial kernel, so distinct
ceiling-states share a residue; single-residue recovery fails. The filtration statement is
Lemma~\ref{lem:cauchy} applied to the residue sequence, and consumes the bridge $\pi_O=T_t$ (M) and
that lemma's hypotheses.
\end{proof}

\begin{conjecture}[Exportability correlates with regime]
Smooth insights have high $\chi_T$; pivot-heavy insights have high $\epsilon_T$. By
Prop.~\ref{prop:chi} (orthogonal case) these are one scalar on the unit sphere; the conjecture is
that this scalar correlates with the measured smooth/pivot label. Falsifiable, not proved (C).
\end{conjecture}

\begin{remark}[What remains around the skeleton]
The skeleton (bounded rank $\Rightarrow$ non-determination $\Rightarrow$ Cauchy-filtration
recovery, Prop.~\ref{prop:skeleton}) is P consuming M; the conjecture above is its only empirical
residue.
\[
  \text{rich insight} \to \text{bounded residue} \to \text{witnessed refinement} \to \text{stable completion}.
\]
PyLantern (\S\ref{sec:pylantern}) is the floor instrument; it records residues, not the private
ceiling.
\end{remark}

\begin{remark}[Weak Born analogy only, firewalled]
Projecting a rich state onto a bounded record is Born-\emph{like} only in the weak sense that a
high-dimensional configuration is compressed into what an observer can register. No claim about literal
quantum measurement is intended. This weak analogy is kept distinct from, and does not assume, the
strong route (a genuine Born derivation via \S\ref{sec:void} branch-invariance supplying Gleason's
non-contextuality hypothesis, with category-theoretic typicality). That route is a separate, stronger,
\emph{unclaimed} target.
\end{remark}

\begin{definition}[Representation-independence obstruction]\label{def:rio}
A \emph{representation-independence obstruction} (RIO) is a triple: a structure-preserving
projection/quotient faithful on a distinguished sub-object, a complement on which it is lossy, and a
non-contextuality (representation-independence) hypothesis asserting the complement
inessential---which the projection alone does not supply.
\end{definition}

\begin{remark}[One obstruction, five axes]
Formalize the shape once (Def.~\ref{def:rio}). Five gaps in this program are RIO-instances:
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{0.26\textwidth} p{0.30\textwidth} p{0.36\textwidth}@{}}
\toprule
\textbf{Axis} & \textbf{Low end (projection exists)} & \textbf{High end (complement inessential)}\\
\midrule
Forcing correspondence & structural transduction (KJ clauses transfer) & identity with Cohen forcing (poset representation-independent)\\
Born & compression onto bounded register & $|\psi|^2$ forced by Gleason non-contextuality\\
Deparameterization (\S\ref{sec:deparam}) & Reading~A: floor fixed by family $R$ & Reading~B: floor fixed by \emph{every} reduction family\\
Inter-observer (\S\ref{sec:void}) & commensurability at the void (Lem~\ref{lem:void}) & contingent claim transported across branches\\
Observer closure (zombie, Prop~\ref{prop:zombie}) & behavioral frame validates the trace & behavior certifies observerhood under unsampled frames\\
\bottomrule
\end{tabular}
\end{center}
Each low end is functorial and cheap; each high end is a uniqueness/non-contextuality theorem strong
enough to fail generically. \emph{Why coupled here:} \S\ref{sec:void} branch-invariance is the common
upstream input, consumed at two strengths---as commensurability it is proved (Lem~\ref{lem:void}); as
the non-contextuality hypothesis a Born derivation would need, it is open. \emph{Against overclaiming:}
the five share a \emph{shape}; the first four also share one upstream hypothesis (\S\ref{sec:void}
branch-invariance), while observer closure has a distinct upstream (projection orthogonality,
Prop~\ref{prop:chi}); they are not the same \emph{object}, and discharging one high end does not
discharge another. The claim earned is schema-level: all five are RIO-instances. Reading the table as a transport would reintroduce the
identity-for-transduction inflation the note refuses.
\end{remark}

\section{Deparameterization and the judgment residue}\label{sec:deparam}

Each preceding section replaced a free parameter by an object the remaining structure already forces.
This section makes that operation the object, and is explicit about how little is proved.

\medskip\noindent\textbf{The operator.} A \emph{specification} $s$ fixes values for parameters
$\Theta(s)$ of the apparatus. Order specifications by parameter removal,
$s \preceq s' :\Longleftrightarrow \Theta(s')\subseteq\Theta(s)$ and $s'$ agrees with $s$ on
$\Theta(s')$; the top $\mathbf{1}$ is the fully deparameterized apparatus, $\Theta=\emptyset$. Let
$R(s)$ be $s$ with every parameter \emph{forced by the rest} removed; $R$ is inflationary,
$R(s)\succeq s$.

\begin{lemma}[Deparameterization closure]\label{lem:bw}
If the specification poset $\Sspec$ is chain-complete, $R$ has a fixed point above every $s$
(Bourbaki--Witt \cite{bourbaki-witt}), and the iterates reach $R^\infty(s_0)$, the most reduced
specification forced from $s_0$. The internal Bourbaki--Witt principle \cite{bw-topos} places this in
the same topos as the Kripke--Joyal semantics of \S\ref{sec:schema}.
\end{lemma}

\begin{proof}
$R$ is inflationary on $\Sspec$ by construction; Bourbaki--Witt applies verbatim to inflationary
maps on chain-complete posets, yielding a fixed point above each $s$ as the stationary value of the
transfinite iteration. The internalization is \cite{bw-topos}. Chain-completeness of $\Sspec$ is the
consumed hypothesis, recorded in the ledger.
\end{proof}

\begin{proposition}[Residue dichotomy]\label{prop:dich}
$R^\infty(s_0)$ is either the top $\mathbf{1}$ or a fixed point $\prec\mathbf{1}$: a \emph{residue}
carrying at least one parameter the remaining structure does not force. This is relative to the family
$R$ (``Reading~A'').
\end{proposition}

\begin{proof}
$R^\infty(s_0)$ is a fixed point of $R$ (Lemma~\ref{lem:bw}). If it is not $\mathbf{1}$, it retains
a nonempty parameter set $\Theta$; fixedness means $R$ removes none of them, and $R$ removes exactly
the parameters forced by the remaining structure, so at least one retained parameter is unforced.
Relativity to the family $R$ is definitional.
\end{proof}

\medskip\noindent\textbf{The bridge to the void (replacing $\mathbf{1}=\bot$).} $\mathbf{1}$ lives in
$\Sspec$; the void $\mathbb{1}$ lives in the condition poset $P_H$. Object-level equality would be a
category error. Instead posit a monotone bridge map (modeling postulate M-Bridge)
\[
  F:\Sspec\to P_H,\qquad F(\mathbf{1})=\mathbb{1},
\]
transporting the terminal specification to the void condition: \emph{the equality is not object-level;
it is functorial transport of the terminal specification to the void.} Under M-Bridge, the fully
deparameterized apparatus maps to the self-determined indexical origin.

\begin{theoremtarget}[Structural irreducibility]
Under M-Bridge, the universal reduction closure is either $\mathbf{1}$---in which case, via
$F(\mathbf{1})=\mathbb{1}$, the irreducible term is not a residual parameter but the indexical origin
itself, irreducible structurally (the fixed point of removal, and by Lemma~\ref{lem:void} the unique
zeroth-order point no first-order operation reaches, while $R$ is first-order)---or a floor
$s_*\prec\mathbf{1}$ fixed by every reduction, a genuine Reading-B residue distinct from the origin.
Deciding which is the open problem.
\end{theoremtarget}

\begin{remark}[One floor, three charts]
The deparameterization floor, the smooth/pivot boundary, and the reconstruction floor / $K$ parameter
of a companion formalism are \emph{conjecturally} the same object: a floor either real or the artifact
of a bounded reconstructor. What is established is that they share a \emph{shape} (an RIO), not that
they are one object; asserting identity would itself be a Reading-B claim. That the same fork recurs
across independent formalisms is evidence it is an invariant of the program, not an artifact of one.
That companion $K$ is the observer-bounded commutator $K_{\mathcal O}$ (Riemann $=[\nabla,\nabla]$), of
which gravity and the past light cone are two $\Omega_{\mathcal O}$ faces; the identification is
structural, not numerical, developed in the observer-bundle companion and firewalled here
(\S\ref{sec:appendix}), an unclaimed \textbf{O} target.
\end{remark}

\begin{conjecture}[Residue $=$ contract break]
The deparameterization residue $s_*\prec\mathbf{1}$, when it exists, coincides with the boundary of
the smooth regime: the retained parameter is exactly the one whose forcing would require crossing a
break of the displacement contract (Assumption~\ref{asm:smooth}). The measurable form uses the
contact ratio and the label-free router of the companion instrumentation (external to this note);
status C, falsifiable under audit-observer measurement.
\end{conjecture}

\section{Executable interpretation: PyLantern as witness layer}\label{sec:pylantern}

PyLantern is the executable witness layer, \emph{not} an autonomous closure engine. It is a calibrated
structural transducer along
\[
  \text{symbolic claim} \to \text{bounded observer state} \to \text{admissible refinement}
  \to \text{witness/certificate} \to \text{human-gated decision},
\]
under one invariant: \emph{witness and request access only; no direct autonomous loop closure}.
In the language of \S\ref{sec:site}, PyLantern operates strictly inside the witnessed fragment
$\Jwit$: it verifies whether a refinement path is witnessed, bounded, logged, and admissible (a finite
$\Jwit$-witness), and must \emph{not} become the generic filter that autonomously executes every
$\Jadm$-cover. Meeting requirements is the mathematical content of completion in $\Jadm$, not a license
for a running system to enact them. This is precisely the containment $\Jwit\subseteq\Jadm$: the
generated site supplies semantics; the witnessed fragment supplies executable discipline.

PyLantern is also a protocol discipline for exporting residues into the interface medium: it
constrains $E_A$ to carry witnesses, provenance, uncertainty, and consent gates, and constrains $D_B$
so received artifacts are treated as reconstructions (Thm.~\ref{thm:nonid}), not identity with the
sender's ceiling-state.

\section{Chalked as the canonical witnessed medium}\label{sec:chalked}

Chalked is not application fluff; it is the canonical model of the interface layer.

\begin{quote}
Chalked implements the witnessed fragment $\Jwit$ of the generated admissible site: a consent-gated
shared witness surface where bounded observers exchange floor-stable residues under explicit protocol.
\end{quote}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{0.34\textwidth} p{0.60\textwidth}@{}}
\toprule
\textbf{Chalked} & \textbf{Interface layer} \\
\midrule
board & medium $M$ \\
mark & encoded residue $E_A(T_Ax_A)$ \\
reading a mark & reconstruction $D_B(m)$ in the reader's surface \\
chalk color & protocol / contract selector \\
eraser & revision authority \\
blank / void & mapped ignorance (depth-$0$, $\mathbb{1}$) \\
certificate & witness attachment \\
permission gate & non-actuation boundary \\
\bottomrule
\end{tabular}
\end{center}

\begin{proposition}[Chalked correctness invariant]\label{prop:chalked}
A Chalked mark is valid only if it preserves provenance, permission, residue type, and revision
authority, while refusing identity with the originating observer-state: it is $E_A(T_Ax_A)$, never
$x_A$. Equivalently, \emph{a valid mark carries residue, not ceiling; reconstruction, not possession}
(Thm.~\ref{thm:nonid}).
\end{proposition}

Chalked is the smallest object on which the Factoring Theorem (Thm.~\ref{thm:factor}) and the
Non-identity Theorem (Thm.~\ref{thm:nonid}) are simultaneously realized and auditable---hence the
canonical, not incidental, model.

\section{Non-actuation boundary}

Proving the semantics does not require, and does not authorize, actuating the system it describes.
\emph{No proof may require the system to close the loop it claims to make safe.} The executable
interpretation is restricted to shadow execution, dry-runs, bounded witnesses, reproducible traces,
and human-gated transitions ($\Jwit$ only). Quantum resources, if present, are queried as bounded
witnesses or simulators, not autonomous optimizers or actuators. The Cauchy-generic completion
(Lemma~\ref{lem:cauchy}) is a statement about the existence of a limit in $\overline{P}_O$; it is not
an instruction for any process to traverse the trajectory autonomously.

\section{Status ledger}\label{sec:ledger}

Every load-bearing claim carries exactly one status: \textbf{P}~= proved relative to stated
hypotheses; \textbf{D}~= definitional architecture (no proof obligation beyond well-formedness);
\textbf{S}~= specification / implementation-conformance contract; \textbf{M}~= modeling postulate;
\textbf{C}~= measurable conjecture or contract; \textbf{O}~= open. A \textbf{P} result is only as
strong as the \textbf{M}/\textbf{C} it consumes, and the Consumes column lists the \emph{transitive}
closure of named debts, not only the direct ones (the accompanying verification suite recomputes
this column mechanically). The site bound, persistence/consistency, deciding density,
Rasiowa--Sikorski, the propositional Truth Lemma, the margin path form (with a formal refutation of
the v14 per-step induction), the Cauchy--Forcing completion, order--metric compatibility,
non-identity of transport, the exportability identity, and observer closure as projection residue
(v16) are machine-verified in Lean~4 (core
kernel + mathlib layer) in that suite. One debt is \emph{ambient} and stated here once rather than per row:
every result below lives on $\PHeta$, whose definition (Def.~\ref{def:refine}) consumes the
smoothness contract (Assumption~\ref{asm:smooth}); rows list the contract only where it enters
\emph{beyond} that ambient dependence.

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{0.40\textwidth} c p{0.42\textwidth}@{}}
\toprule
\textbf{Claim} & \textbf{Status} & \textbf{Consumes / conditional on}\\
\midrule
Generated site $\Jadm=\langle B\rangle$ (Def~\ref{def:site}) & D & existence is P (proof of Lem~\ref{lem:sitebound})\\
Site bound $\Jadm\subseteq\Jnn$ (Lem~\ref{lem:sitebound}) & P & admissible density of generators\\
Witnessed fragment $\Jwit\subseteq\Jadm$ (Def~\ref{def:wit}) & D & finite-witness definition; discipline\\
Smoothness contract (Asm~\ref{asm:smooth}) & C & measured, substrate-relative\\
Persistence \& consistency (Lem~\ref{lem:pers}) & P & pullback stability; M-Pers only via well-formedness ($A_\varphi$ a sieve)\\
Channel-Margin, path form (Lem~\ref{lem:margin}) & P & Asm~\ref{asm:smooth} (C): Lipschitz + budget\\
Separativity $=$ certificate disagreement (Lem~\ref{lem:nonatom}) & P & stated for $\forcesnn$; M-Cvx; M-Pers (certificate persistence)\\
Boundedness $\Rightarrow$ countability (Lem~\ref{lem:bdd}) & P & observer bound: M-Bound\\
Decision Reachability (Lem~\ref{lem:reach}) & P & Asm~\ref{asm:smooth} (C); M-Pers (via Def~\ref{def:force}); item 8 of queue\\
Cauchy--Forcing completion (Lem~\ref{lem:cauchy}) & P & $V$: M; descent (b): C; $d_O$ comparability (c): M (\emph{Principia})\\
Order--Metric compatibility, atomic (Lem~\ref{lem:ordmet}) & P & M-Pers; stabilized region closed: M-Cl\\
Atomic Truth Lemma (Lem~\ref{lem:atomic}) & P & M-Pers; Asm~\ref{asm:smooth} (C) and M-Bound via $\mathcal{D}_p^\ast$\\
Deciding sets order-dense (Lem~\ref{lem:dec}) & P & M-Pers (via Def~\ref{def:force}: $A_\varphi$ a sieve)\\
Genericity decides (Lem~\ref{lem:bivalence}) & P & Lem~\ref{lem:reach} (hence Asm~\ref{asm:smooth}: C); M-Bound; M-Pers\\
Propositional Truth Lemma (Thm~\ref{thm:prop}) & P & M-Pers; Asm~\ref{asm:smooth} (C, via reach); M-Bound; calibration queue: \textbf{O}\\
Torsion $=$ $\neg$-clause, atomic (Rem.~\ref{rem:torsion}) & P & atomic only; M-Pers (via $A_\varphi$); compound case \textbf{O} (finite countermodel)\\
Void is unique shared point (Lem~\ref{lem:void}) & P & wedge construction\\
Forced-to-zeroth-order (Prop~\ref{prop:zeroth}) & P & refinement depth; Lem~\ref{lem:void}\\
\;\;$\hookrightarrow$ diagnostic-register id. & M & modeling bridge, not the order fact\\
Transfer (Def~\ref{def:transfer}) & D & interface architecture\\
Factoring through $M$ (Thm~\ref{thm:factor}) & D & Lem~\ref{lem:void} + Def~\ref{def:transfer}; exclusion content is P\\
Non-identity of transport (Thm~\ref{thm:nonid}) & P & $T_A$ idempotent; equality condition named\\
Observer closure $=\epsilon_T$ residue (Prop~\ref{prop:zombie}) & P & bridge M-Zomb; Thm~\ref{thm:nonid}(i), Prop~\ref{prop:chi}, Rem~\ref{rem:torsion}\\
Interface maps $E_A,D_B$ semantics & M & protocol discipline\\
Chalked correctness invariant (Prop~\ref{prop:chalked}) & S & implements Thm~\ref{thm:nonid} discipline\\
\midrule
\multicolumn{3}{@{}l}{\textit{Interpretive bridges / appendices}}\\
Eureka-Tub skeleton (Prop~\ref{prop:skeleton}) & P & $\pi_O=T_t$: M; Lem~\ref{lem:cauchy} hypotheses\\
Exportability identity $\chi^2+\epsilon^2=1$ (Prop~\ref{prop:chi}) & P & $T_t$ orthogonal; oblique case: triangle bound only\\
$\chi/\epsilon \leftrightarrow$ smooth/pivot correlation & C & measured\\
Weak Born (compression) & M & analogy; strong route (Gleason): \textbf{O}\\
RIO schema (Def~\ref{def:rio}; one obstruction, five axes) & D & schema-level only; not object identity\\
Curvature bridge $K=K_{\mathcal O}$ (gravity/light-cone $\Omega_{\mathcal O}$ faces) & O & structural, not numerical; observer-bundle companion; firewalled\\
Deparameterization closure (Lem~\ref{lem:bw}) & P & chain-completeness (Bourbaki--Witt)\\
Residue dichotomy (Prop~\ref{prop:dich}) & P & relative to family $R$ (Reading~A)\\
Bridge $F(\mathbf{1})=\mathbb{1}$ & M & M-Bridge (functorial transport, not equality)\\
Residue $=$ contract break (Conjecture) & C & contact ratio + label-free router\\
Residue is the origin (Reading~B) & O & the program's single open problem\\
Loopback as self-reference icon & M & iconographic, nonessential to proof\\
\bottomrule
\end{tabular}
\end{center}

\noindent \textbf{(M-Pers) is the highest-leverage postulate}, with a sharper blast radius than v14
claimed: clause-level persistence of the forcing relations (Lemma~\ref{lem:pers}) consumes it only
through the well-formedness of the atomic relation---the persistence \emph{argument} is pullback
stability of the topology, and survives M-Pers removal at the dense reading (mutation-tested in
the verification suite). What M-Pers substantively funds is (i) the sieve-hood of the stability sets
$A_\varphi$ (well-formedness of Def.~\ref{def:force}), (ii) the $(\Rightarrow)$ direction of the
Atomic Truth Lemma, and (iii) the Order--Metric Compatibility Lemma; through (ii) it reaches the
Propositional Truth Lemma. If it fails, the truth lemmas fail while the forcing calculus itself
stands. It remains the first target for derivation from the apparatus. No entry is an unconditional
\textbf{P} about the world; each is a \textbf{P} about the mathematics plus a named
\textbf{M}/\textbf{C}/\textbf{O} debt to the substrate.

\section{Boundary and caution}

This is structural correspondence, not identity with Cohen forcing. The generated-site framing
(\S\ref{sec:site}) makes the schema a theorem---forcing over $(\PHeta,\Jadm)$, with $\Jadm\subseteq\Jnn$
free---rather than an analogy, while keeping raw operational admissibility from bearing a brittle
closure claim. Channel viability with persistence (Lem~\ref{lem:margin}), certificate separativity
(Lem~\ref{lem:nonatom}), generic existence (Lem~\ref{lem:bdd}), Cauchy-generic completion with
order--metric compatibility (Lems~\ref{lem:cauchy}, \ref{lem:ordmet}), atomic soundness
(Lem~\ref{lem:atomic}), and propositional soundness (Thm~\ref{thm:prop}) are discharged relative to the
ledger. Two layers remain genuinely open. \emph{Quantifiers}: $\exists,\forall$ need covering/fullness
clauses and witness construction (candidate: obstruction-dual generators 3a--3d; not verified).
\emph{Calibration}: the finite theorem queue of \S\ref{sec:calibration}, of which Decision
Reachability and the per-connective realization lemmas are the load-bearing items---an implementation
check, the transduction-not-identity discipline one level up.

\section{Suggested paragraph for \emph{The Hypothesis Surface}}

\begin{quote}
The surface carries an order: one state refines another when it keeps the earlier commitments and
resolves more, under the admissible moves the budget allows. The admissible dense obligations the
observer emits \emph{generate} a site; on it a forcing relation appears---a state forces a proposition
when the stabilizing refinements cover it---and it is tempting to read it as Cohen forcing relocated.
We hold the weaker claim: both instantiate the same site semantics, the Kripke--Joyal interpretation
that makes Cohen's construction work, differing only in which requirements generate the covers. Two
conditions the reading demands are present rather than imposed: the channel keeps a positive margin,
$\lambda_{\min}(\tilde\Gamma)\ge\eta/2$, by a path-budget perturbation estimate (the metric nondegeneracy
that makes refinements exist, while incompatibility is carried by certificate disagreement); and the
generic completion exists because a bounded observer holds only countably many requirements, the
hypothesis Rasiowa--Sikorski consumes. Soundness holds for atomic propositions and, conditional on a
finite calibration queue, for the connectives; what remains is the quantifier layer and the check that
the surface operations realize the clauses.
\end{quote}

\section*{Closing invariant}

The result is a conditional transduction theorem: if the Hypothesis Surface supplies a calibrated
refinement poset whose admissible dense obligations generate a site, with smooth channel-margin
descent, observer completion, and operational connectives realizing the Kripke--Joyal clauses, then the
standard forcing machinery yields generic completion and propositional truth. \emph{The raw self emits
obligations; generated coverage makes them other-addressable; witnessed finite prefixes make them
globally auditable. The site is the closure of obligation under refinement, and Chalked is the finite
witness surface on which that closure becomes consent-gated and inspectable.} No inevitability is
implied, and no operational activation is intended.

\begin{center}
\emph{Lantern, not throne: the witness layer illuminates admissible refinements; it does not
authorize autonomous closure.}
\end{center}

\appendix
\section{Iconography, firewalls, and scope}\label{sec:appendix}

\noindent\textbf{Loopback.} \texttt{127.0.0.1} is retained as iconography for indexical self-reference
(universal form, local referent); it is nonessential to every proof and bridges no observers.

\noindent\textbf{Weak Born firewall.} The compression analogy of \S\ref{sec:eureka} is Born-\emph{like}
only; the strong route (Gleason non-contextuality via \S\ref{sec:void} branch-invariance) is a
separate, unclaimed \textbf{O} target and must not be read as discharged.

\noindent\textbf{Curvature firewall.} Identifying the companion floor $K$ with the observer-bounded
commutator $K_{\mathcal O}$---gravity and the light cone as $\Omega_{\mathcal O}$ faces---is structural,
not numerical; it is developed in the observer-bundle companion and is an unclaimed \textbf{O} target
here. The apparent torsion/non-torsion $=$ curvature/torsion pun is to be verified, not asserted.

\noindent\textbf{Singularity, myth, authority.} In a venue where ``singularity'' is field vocabulary,
this work argues for a non-actuating interpretation: witness surfaces, bounded refinement, human-gated
completion, not autonomous closure. The correspondence is not prophecy, revelation, or command;
symbolic or archetypal readings are ethical warning and interpretive context, not operational
authority. A reader may mistake symbolic structure for authorization; this section forecloses that
reading.

\begin{thebibliography}{9}
\bibitem{rasiowa-sikorski} H.~Rasiowa and R.~Sikorski. A proof of the completeness theorem of
G\"odel. \emph{Fundamenta Mathematicae} \textbf{37} (1950), 193--200.
\bibitem{gpst} E.~Gunther, M.~Pagano, and P.~S\'anchez Terraf. Formalization of forcing in
Isabelle/ZF. In \emph{Automated Reasoning} (IJCAR 2020), LNCS \textbf{12167}, Springer, 2020.
\bibitem{chow} T.~Y.~Chow. A beginner's guide to forcing. In \emph{Communicating Mathematics},
Contemporary Mathematics \textbf{479}, AMS, 2009, 25--40. arXiv:0712.1320.
\bibitem{maclane-moerdijk} S.~Mac Lane and I.~Moerdijk. \emph{Sheaves in Geometry and Logic: A
First Introduction to Topos Theory}. Springer, 1992.
\bibitem{bourbaki-witt} N.~Bourbaki. Sur le th\'eor\`eme de Zorn. \emph{Archiv der Mathematik}
\textbf{2} (1949/50), 434--437; E.~Witt. Beweisstudien zum Satz von M.~Zorn. \emph{Mathematische
Nachrichten} \textbf{4} (1951), 434--438.
\bibitem{bw-topos} The internal Bourbaki--Witt principle in toposes; arXiv:1201.0340.
\end{thebibliography}

\bigskip
\noindent\rule{\linewidth}{0.4pt}\\[0.3em]
\footnotesize
The collapse geometry ($\delta_{\min}\propto\sqrt{k/(1-\rho(k-1))}$, the cliff at $\rho(k-1)=1$,
the perturbation sensitivity $\sim 1/\lmin$) is from \emph{The Cost of Cacophony} (Tiffany, in
preparation). \emph{Source/author attribution: Paul Tiffany.}

\end{document}
