\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,hyperref}
\usepackage[T1]{fontenc}
\input{generated_book5_counts.tex}

\title{Lean PS: Typed Correspondence, Witness Persistence,\\
and a Verified Kernel for Principia Symbolica Book V}
\author{Working paper scaffold for the AGI-26 correspondence program}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
We describe a machine-checked architecture connecting three formally distinct
domains: Kripke--Joyal forcing under refinement, covariance of physical force
laws, and selected arithmetic and conditional-system claims from
\emph{Principia Symbolica} Book V.  The connection is a shared typed
commutation interface, not an identification of semantics.  The current receipt
contains \VerifiedDeclarationCount\ verified declarations across the analysis
stack.  Of these, \MappedBookFiveDeclarationCount\ declarations map to
\MappedBookFiveAnchorCount\ Book V anchors.  The Book V atlas contains
\BookFiveClaimNodeCount\ theorem-like claims, of which
\OpenBookFiveClaimCount\ remain explicitly unmapped.  We present the proof
boundary, negative controls, and a receipt-driven pedagogy pipeline that refuses
to promote paper prose to kernel fact.
\end{abstract}

\section{The correspondence is an interface}

The common object is a commuting condition.  In the total setting,
equivariance states
\[
  f(g\cdot x)=g\cdot f(x).
\]
In the relational forcing setting, an admissible refinement step is preserved
by the forcing predicate.  The two constructions instantiate one interface but
differ in totality, equality strength, and semantics.  The formal architecture
therefore retains domain, process schema, observer scope, adjudication mode,
and transport-loss class as independent coordinates.

This separation blocks two invalid inferences.  First, agreement across
observers does not imply persistence under refinement.  Second, commutation of
a physical square does not reflect membership in the Lorentz group without
nondegeneracy.  Both failures have machine-checked finite or algebraic
countermodels.

\section{Witness preservation and M-Pers}

A witness system assigns certificate types $W(o,p,a)$ to observer $o$,
condition $p$, and atom $a$, together with restriction
\[
  q\preceq p \quad\Longrightarrow\quad W(o,p,a)\to W(o,q,a).
\]
Stability is inhabitation of the certificate type.  Hence local persistence is
derived:
\[
  \operatorname{Stable}(o,p,a)\land q\preceq p
  \Longrightarrow \operatorname{Stable}(o,q,a).
\]
Cross-observer transport is a second structure with a named naturality law.
Material persistence may use both axes, but horizontal agreement alone does
not establish the vertical theorem.

\section{Book V proof families}

\subsection{Balanced memory and reciprocity}

Balanced two-step memory uses
\[
 A=\begin{pmatrix}1&1\\1&0\end{pmatrix}.
\]
Lean checks the eigenpair $(\varphi,1)$, uniqueness of the positive root of
$x^2=x+1$, and convergence of consecutive Fibonacci ratios.  Weighted
reciprocity uses
\[
 T_w=\begin{pmatrix}1&w\\1&0\end{pmatrix},\qquad
 \lambda_w=\frac{1+\sqrt{1+4w}}2.
\]
The kernel proves $\lambda_w^2=\lambda_w+w$, the corresponding eigenvector,
$\lambda_1=\varphi$, sub-golden rates for $0<w<1$, and super-golden rates for
$w>1$.  Claims about ethics, decency, or agency remain interpretations of this
conditional model.

\subsection{Fracture and product spectra}

For a nonzero equal-coordinate plane diagonal, Lean proves
\[
 \frac{\|v\|_1}{\|v\|_2}=\sqrt2,
 \qquad
 \frac{\|v\|_1}{\|v\|_\infty}=2.
\]
Separate countermodels show that fixing the memory regime does not determine
the norm regime and fixing the norm regime does not determine memory.  Thus the
symbolic-geometric interface is represented by typed product coordinates
rather than a single sovereign classification.  The finite-support kernel now
includes positive finite $p$, the infinity regime, sharp equality cases,
effective support dimension, and finite symbolic-path decoherence.

\subsection{Conditional thermodynamics}

Free energy and viability are definitions.  Conservation, entropy production,
fixed-point equivalence, persistence, and covenant stability are represented by
structures carrying their laws.  Lean proves consequences from those fields;
it does not install Book V's modeling laws as global axioms.  For example,
nonnegative entropy rate follows from the explicit comparison
\[
  \rho_{\mathrm{reflection}}\leq\sigma_{\mathrm{production}},
\]
and persistence follows only after supplying a positive-energy persistence
bridge.

Coherence balance is likewise an energy-budget identity, not zero energy.
When drift and reflection rates agree, Lean proves only that the energy rate
equals the external rate.  Conservation requires the additional closed-budget
hypothesis; a concrete compiled countermodel refutes conservation from balance
alone.

\section{Receipt-driven pedagogy}

The build emits a theorem receipt containing declaration name, source module,
axiom dependencies, and a no-\texttt{sorry} gate.  A second projection maps
Principia anchors to exact, conditional, partial, or definitional coverage.
Matt consumes these claims; Ellie composes a semantic lesson; the transcript
gate rejects ungrounded numbers, status words, and guarded universalizations.
The same semantic program is suitable input for textbook, slides, audio, or
video renderers.  Rendering is downstream of proof and does not strengthen it.

\section{Open program and workshop ingestion}

The immediate proof queue is:
\begin{enumerate}
  \item deployed Surface witness types and a construction of M-Pers for actual
        transition rules;
  \item trajectory-response hypotheses for Book V adaptation, kept separate
        from life classification and metabolic regulation;
  \item spectral hypotheses sufficient for the MAD--MAP--MAS classification;
  \item compact or measurable path-space extensions when later claims require
        more than the finite symbolic-decoherence kernel;
  \item functorial witness restriction and observer transport.
\end{enumerate}

A future AGI-26 workshop paper can be ingested by mapping each statement to
this receipt discipline: definition, proved theorem, conditional theorem,
measurable conjecture, modeling postulate, or open debt.  Shared diagrams do
not merge their domains.  The weakest adequate composition remains the
governing rule.
\section{Observer lowering and complementary half-bridges}

The quantum correspondence is directional.  Its two Gleason-facing halves are
\[
\begin{aligned}
\text{normalized source} &\longrightarrow \text{Hermitian density}
  \longrightarrow \text{resolved readout},\\
\text{certified readout laws} &\longrightarrow
  \text{Hermitian or sesquilinear representation}.
\end{aligned}
\]
The lower line represents observer-visible values; it does not invert the upper
line.  Global phase supplies a kernel-certified collision at the seam, and the
frame and resolution countermodels block unconditional reverse reconstruction.
Finite partial trace is separately certified as a reduction preserving trace and
represented local expectations.

The temporal arrow is independently formal.  The Cacophony-facing kernels certify
finite-support norm fracture and a strict diagonal path-cost gap; ordered stage costs
telescope; and a minimizing JKO step pays its transport displacement from free-energy
decrease.  Summable stage bounds in a complete observer metric, or positive
Lyapunov descent paying each increment, certify convergence.  The interpretive step
is narrower and functional: mapping physical quantum decoherence and noise onto
this general directed transport-and-loss geometry.  Like a weather map, the
correspondence is operational and testable without asserting a kernel identity between
the mapped domains.

\section{Reproducibility boundary}

The paper compiles only after
\texttt{book5\_lean\_coverage.py} generates the count macros above.
The JSON coverage packet lists every mapped anchor and every unmapped
claim-bearing Book V node.  The Lean root prints the axiom closure of each
headline declaration.  Numerical Lorentz and Plomp--Levelt witnesses remain
separate conformance artifacts.  Poems and interpretive prose are preserved as
authored content and are not represented as proof obligations.

\end{document}
